---
title: "CTA-ED Exercise 5: Unsupervised learning (topic models)"
author: "Nayah and Min"
date: "12/03/2024"
output: html_document
---
## Exercises

## Setup 

Before proceeding, we'll load the packages we will need for this tutorial.

```{r, message=F}
library(tidyverse) # loads dplyr, ggplot2, and others
library(stringr) # to handle text elements
library(tidytext) # includes set of functions useful for manipulating text
library(topicmodels) # to estimate topic models
library(gutenbergr) # to get text data
library(scales)
library(tm)
library(ggthemes) # to make your plots look nice
library(readr)
library(quanteda)
library(quanteda.textmodels)
```

```{r}
#install_package(devtools)
devtools::install_github("matthewjdenny/preText")
library(preText)
```

## Question 1: War and Peace

1. Choose another book or set of books from Project Gutenberg
We choose War and Peace, a novel composed of 15 books + epilogue.  
```{r choosebook}
tolstoy <- gutenberg_download(c(2600),
                            meta_fields = "author")

```

In order to create a dfm, we take out the lengthy table of contents, and divide it up by book number, before passing the resulting data to a dataframe.

```{r dfm}

# first we need to take out lines 1-796, which are table of contents, and the irrelevant info.
tolstoy_edit<- tolstoy %>%
  select(-gutenberg_id,-author)%>%
  mutate(row.id=1:nrow(tolstoy))%>%
  filter(row.id>797)

Book_Headings <- tolstoy_edit[grep("BOOK", tolstoy_edit$text), ]
Book_Headings[16:30,]
book.name<-c(Book_Headings$text,"EPILOGUE")#  create book name vector

#We make a vector with the rirght length of each book/chapter id.
row_chapt<-c(Book_Headings$row.id,nrow(tolstoy_edit)) #the row.ids
prior<-c(0,row_chapt[1:15]) #number of rows already accounted for
repetitions<-row_chapt-prior #how many times to repeat

book.id<-rep(book.name,times=repetitions) #create vector for df
nrow(tolstoy_edit)==length(book.id)#check it worked lengthwise, true

tolstoy_edit2<-cbind(tolstoy_edit,book.id) #we bind this to the text dataframe.



tolstoy_words <- tolstoy_edit2 %>% # we now run normal preprocessing to create a dtm.
  unnest_tokens(word, text) %>%
  filter(!is.na(word)) %>%
  count(book.id, word, sort = TRUE) %>%
  ungroup() %>%
  anti_join(stop_words) 


tolstoy_dtm <-tolstoy_words %>% #create dtm
  cast_dtm(book.id, word, n)

tm::inspect(tolstoy_dtm) #look at dtm
```


## Question 2: Own topic model
2. Run your own topic model on these books, changing the k of topics, and evaluating accuracy.

Some of the plausible topic distinctions in War and Peace would be: distinguishing by books (16), or distinguishing by family subplots (5), or between narrative and philosophy (2). We will run them seperately, staring with the lowest k. Vis

```{r modelk2}
tolstoy_lda_2 <- LDA(tolstoy_dtm, k = 2, control = list(seed = 1209)) #run model

tolstoy_lda_2 #view model

tolstoy_topics_2 <- tidy(tolstoy_lda_2, matrix = "beta") #extract the per-topic-per-word probabilities

head(tolstoy_topics_2, n = 10) #look at topics.

```

We can also run it with family subplots.
```{r modelk5}
tolstoy_lda_5 <- LDA(tolstoy_dtm, k = 5, control = list(seed = 1920)) #run model

tolstoy_lda_5 #view model

tolstoy_topics_5<- tidy(tolstoy_lda_5, matrix = "beta")

head(tolstoy_topics_5, n = 10) #look at topics.

tolstoy_top_terms_5 <- tolstoy_topics_5 %>% #arrange topics
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

tolstoy_top_terms_5 %>% # plot topics
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 4) +
  scale_y_reordered()

```

We are also looking at 10. 
```{r modelk10}
tolstoy_lda_10 <- LDA(tolstoy_dtm, k = 10, control = list(seed = 1230)) #run model

tolstoy_lda_10 #view model

tolstoy_topics_10<- tidy(tolstoy_lda_10, matrix = "beta")

head(tolstoy_topics_10, n = 10) #look at topics.

tolstoy_top_terms_10 <- tolstoy_topics_10 %>% #arrange topics
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

tolstoy_top_terms_10 %>% # plot topics
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 4) +
  scale_y_reordered()

```

And finally, 16. 
```{r modelk16}
tolstoy_lda_16 <- LDA(tolstoy_dtm, k = 16, control = list(seed = 1203)) #run model

tolstoy_lda_16 #view model

tolstoy_topics_16<- tidy(tolstoy_lda_16, matrix = "beta")

head(tolstoy_topics_16, n = 10) #look at topics.

tolstoy_top_terms_16 <- tolstoy_topics_16 %>% #arrange topics
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

tolstoy_top_terms_16 %>% # plot topics
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 4) +
  scale_y_reordered()

```
These all provide some form of topic distinction, with some topics easy to distinguish: such as army vs love vs philosophical topics, or the key personages in each of them. However, in order to formally evaluate them, we need to look closer. 


### Split into chapter documents

In the below, we first separate the volumes into chapters, then we repeat the same procedure as above. The only difference now is that instead of one document per book, we have one document per chapter, increasing sparsity

```{r}

# Make a check if there are any NAs in "tolstoy_edit2"
any(is.na(tolstoy_edit2)) 

# Divide into documents, each representing one chapter
tolstoy_chapter <- tolstoy_edit2 %>%
  group_by(book.id) %>%
  mutate(chapter = cumsum(str_detect(text, regex("^chapter ", ignore_case = TRUE)))) %>%
  ungroup() %>%
  filter(chapter > 0) %>%
  unite(document, book.id, chapter)

# Split into words
tolstoy_chapter_word <- tolstoy_chapter %>%
  unnest_tokens(word, text)

# Find document-word counts
tolstoy_word_counts <- tolstoy_chapter_word %>%
  anti_join(stop_words) %>%
  count(document, word, sort = TRUE) %>%
  ungroup()

tolstoy_word_counts

# Cast into DTM format for LDA analysis

tolstoy_chapters_dtm <- tolstoy_word_counts %>%
  cast_dtm(document, word, n)

tm::inspect(tolstoy_chapters_dtm)

```

We then re-estimate the topic model with this new DocumentTermMatrix object, specifying k equal to 16. This will enable us to evaluate whether a topic model is able to generatively assign to volume with accuracy.

```{r}
tolstoy_chapters_lda <- LDA(tolstoy_chapters_dtm, k = 16, control = list(seed = 1249))
```

After this, it is worth looking at another output of the latent dirichlet allocation procedure. The γ probability represents the per-document-per-topic probability or, in other words, the probability that a given document (here: chapter) belongs to a particular topic (and here, we are assuming these topics represent volumes).

The gamma values are therefore the estimated proportion of words within a given chapter allocated to a given volume. 

```{r}

tolstoy_chapters_gamma <- tidy(tolstoy_chapters_lda, matrix = "gamma")
tolstoy_chapters_gamma

```

### Examine consensus

Now that we have these topic probabilities, we can see how well our unsupervised learning did at distinguishing the books generatively just from the words contained in each chapter.

```{r}

# First separate the document name into title and chapter

tolstoy_chapters_gamma <- tolstoy_chapters_gamma %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE)

tolstoy_chapter_classifications <- tolstoy_chapters_gamma %>%
  group_by(title, chapter) %>%
  top_n(1, gamma) %>%
  ungroup()

tolstoy_book_topics <- tolstoy_chapter_classifications %>%
  count(title, topic) %>%
  group_by(title) %>%
  top_n(1, n) %>%
  ungroup() %>%
  transmute(consensus = title, topic)

tolstoy_chapter_classifications %>%
  inner_join(tolstoy_book_topics, by = "topic") %>%
  filter(title != consensus)

# Look document-word pairs were to see which words in each documents were assigned
# to a given topic

assignments <- augment(tolstoy_chapters_lda, data = tolstoy_chapters_dtm)
assignments

assignments <- assignments %>%
  separate(document, c("title", "chapter"), sep = "_", convert = TRUE) %>%
  inner_join(tolstoy_book_topics, by = c(".topic" = "topic"))

assignments %>%
  count(title, consensus, wt = count) %>%
  group_by(title) %>%
  mutate(percent = n / sum(n)) %>%
  ggplot(aes(consensus, title, fill = percent)) +
  geom_tile() +
  scale_fill_gradient2(high = "red", label = percent_format()) +
  geom_text(aes(x = consensus, y = title, label = scales::percent(percent))) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank()) +
  labs(x = "Book words assigned to",
       y = "Book words came from",
       fill = "% of assignments")
?geom_text
```

Not bad! We see that the diagonal, with few exceptions, shows the highest correlation, implying that the model did a good job.


## Question 3: Validation using `preText`

In this section, we'll be using the `preText` package mentioned in @denny_text_2018 to see the impact of different pre-processing choices on our text.

First we need to reformat our text into a `quanteda` corpus object. 

```{r}
# load in corpus of the text data.
corp_tolstoy <- corpus(tolstoy_edit2, text_field = "text")
# use first 10 documents for example
documents_tolstoy <- corp_tolstoy[sample(1:30000,1000)]
# take a look at the document names
print(names(documents_tolstoy[1:10]))
```

And now we are ready to preprocess in different ways. Here, we are including n-grams so we are preprocessing the text in 128 different ways. This takes about ten minutes to run on a machine with 8GB RAM.

```{r}
preprocessed_documents_tolstoy <- factorial_preprocessing(
    documents_tolstoy,
    use_ngrams = TRUE,
    infrequent_term_threshold = 0.2,
    verbose = FALSE)
```

We can then get the results of our pre-processing, comparing the distance between documents that have been processed in different ways. 

```{r,warning = F}
preText_results1 <- preText(
    preprocessed_documents_tolstoy,
    dataset_name = " new text",
    distance_method = "cosine",
    num_comparisons = 20,
    verbose = FALSE)

```

And we can plot these accordingly. 

```{r,warning = F}
preText_score_plot(preText_results1)
```


